New insights into the Weddell Sea ecosystem applying a quantitative network approach

### 2023-12-18 Manuscript/Manuscript.Rmd
Incorporating editors suggs

The high Antarctic Weddell Sea shelf is situated between 74 and 78ºS, stretching approximately 450 km from East to West (Figure 1). Water depth varies between 200 and 500 meters, and shallower areas are covered by continental ice, which forms the coastline along the eastern and southern parts of the Weddell Sea. The shelf area contains a complex three-dimensional benthic habitat with large benthic biomasses, intermediate to high diversity in comparison to benthic boreal communities and a spatially patchy distribution of organisms [@Dayton1990; @Teixido2002].

@@

The Weddell Sea food web was retrieved from the GlobAL daTabasE of traits and food Web Architecture (GATEWAy, version 1.0) of the German Centre for Integrative Biodiversity Research (iDiv) Halle-Jena-Leipzig [@Brose2018]. In addition to predator-prey interactions, the database contains information on other biological data such as mean body mass and movement type for each species in the food web. Furthermore, it incorporates information about the interaction itself, such as the dimension of the predator search space (2 or 3 dimensions). In its current form, the Weddell Sea food web comprises 490 species and 16041 predator-prey interactions and thus constitutes one of the most resolved food webs constructed to date [@Jacob2011].

Complex empirical food webs are compilations of species and their potential feeding interactions within a given area and time period. The species occurrence data to construct the Weddell Sea food web (Antarctica) was sampled between 1983 and 2005, and hence the Weddell Sea food web is representative for this time period. The diet composition for each species was observed between 2001 and 2004 through a combination of field observations and stomach content analyses (further details can be found in @Jacob2005). Some species, such as benthic grazers and suspension feeders had poor taxonomic resolution of their prey, therefore laboratory information about their size, behavior, and stable isotope signatures was used to infer their feeding habits [@Jacob2011]. As the data used to construct the Weddell Sea food web is sampled during the summer season due to the inaccessibility of ecological field sampling in Antarctica during the winter, the Weddell Sea food web is representative of the summer season.
Moreover, the species in the food web are representative of adult specimens, and were not categorized into ontogenetic life stages (i.e., larvae, juveniles, and adults). As a result, ontogenetic diet shifts cannot be addressed with this food web. The average body masses of species were either directly measured [@Jacob2005] or taken from published accounts for marine mammals and seabirds [@Brose2006a].
Due to the static nature of the Weddell Sea food web and its low temporal resolution, it is not possible to assess phenological mismatches nor species temporal turnover. Because of these methodological limitations, the nature of the Weddell Sea food web data may likely lead to overestimations of some trophic interactions and network properties (connectance), compared to realised networks at any given point in time and space.
The spatial extent of the food web is described in the Study area section (Figure 1).

@@

Here it is important to note that we observed an important variability when estimating each IS. The interactions that showed the greatest variability, represented by a wide interquartile range, were those where consumer and resource body masses were very small (approximately 1e-15 kg). The frequency distribution of the interquartile range is presented in the Supplementary Material (Figure S1).

@@

Another source of uncertainty stems from the parameters and data used to estimate species IS. While we accounted for variability from estimated resource density and consumer search rate [@Pawar2012], we did not consider potential variability arising from changes in species' body mass. Indeed, one could estimate the interaction strength based on body mass distributions rather than a single empirical value. Future research should consider this source of variability since it could alter food web stability outcomes [@Brose2006a; @Allesina2015; @Gross2009; @Landi2018]. Concerning body mass, the model utilized in this study exhibits sensitivity in estimating interaction strengths (IS) among species with exceedingly small body masses (approximately 1e-15  kg), as evidenced by a broad interquartile range. Additionally, some of the observed results may stem from the fact that the applied methodology does not incorporate empirically measured biomasses for species. For example, we would have expected a relatively high impact on community stability when removing the Antarctic krill Euphausia superba, as it is one of the most important prey species in the food web based on field observations [@Atkinson2019]. Moreover, it is one of the species most affected by warming and its population has already declined as a result of productivity changes and increased access by predators [@Atkinson2004; @Kawaguchi2009]. Although our results did not show a significant impact on the stability of the Weddell Sea food web when krill went extinct, we were able to frame it as a key species in terms of its weighted and unweighted characteristics due to its high mean IS, mid trophic level and high number of interactions.

>>>>

The high Antarctic Weddell Sea shelf, our study area, is situated between 74 and 78ºS, stretching approximately 450 km from East to West (Figure 1). Water depth varies between 200 and 500 meters, and shallower areas are covered by continental ice, which forms the coastline along the eastern and southern parts of the Weddell Sea. The shelf area contains a complex three-dimensional benthic habitat with large benthic biomasses, intermediate to high diversity in comparison to benthic boreal communities and a spatially patchy distribution of organisms [@Dayton1990; @Teixido2002].

@@

The Weddell Sea food web was retrieved from the GlobAL daTabasE of traits and food Web Architecture (GATEWAy, version 1.0) of the German Centre for Integrative Biodiversity Research (iDiv) Halle-Jena-Leipzig [@Brose2018]. In addition to predator-prey interactions, the database contains information on other biological data: mean body mass and length, and movement and metabolic type for each species in the food web. Furthermore, it incorporates information about the interaction itself: type (predacious, herbivorous and detritivorous) and dimension (2 or 3 dimensions according to the predator search space). In its current form, the Weddell Sea food web comprises 490 species and 16041 predator-prey interactions and thus constitutes one of the most resolved food webs constructed to date [@Jacob2011].

Complex empirical food webs are compilations of species and their potential feeding interactions within a given area and time period. The species occurrence data to construct the Weddell Sea food web was sampled between 1983 and 2005, and hence the Weddell Sea food web is representative for this time period. The diet composition for each species was observed between 2001 and 2004 through a combination of field observations and stomach content analyses (further details can be found in @Jacob2005). Some species, such as benthic grazers and suspension feeders had poor taxonomic resolution of their prey, therefore laboratory information about their size, behavior, and stable isotope signatures was used to infer their feeding habits [@Jacob2011]. As the data used to construct the Weddell Sea food web is sampled during the summer season due to the inaccessibility of ecological field sampling in Antarctica during the winter, the Weddell Sea food web is representative of the summer season.
Moreover, the species in the food web are representative of adult specimens, and were not categorized into ontogenetic life stages (i.e., larvae, juveniles, and adults). As a result, ontogenetic diet shifts cannot be addressed with this food web. The average body masses of species were either directly measured [@Jacob2005] or taken from published accounts for marine mammals and seabirds [@Brose2006a].
Due to the static nature of the Weddell Sea food web and its low temporal resolution, it is not possible to assess phenological mismatches nor species temporal turnover. Because of these methodological limitations, the nature of the Weddell Sea food web data may likely lead to overestimations of some trophic interactions and network properties (connectance), compared to realised networks at any given point in time and space.
The spatial extent of the food web is described in the Study area section (Figure 1).

@@

Here it is important to note that we observed an important variability when estimating each IS. The interactions that showed the greatest variability, represented by a wide interquartile range, were those where consumer and resource body masses were very small (approximately $1 x 10 ^{-15}$ kg). The frequency distribution of the interquartile range is presented in the Supplementary Material (Figure S1).

@@

Another source of uncertainty stems from the parameters and data used to estimate species IS. While we accounted for variability from estimated resource density and consumer search rate [@Pawar2012], we did not consider potential variability arising from changes in species' body mass. Indeed, one could estimate the interaction strength based on body mass distributions rather than a single empirical value. Future research should consider this source of variability since it could alter food web stability outcomes [@Brose2006a; @Allesina2015; @Gross2009; @Landi2018]. Concerning body mass, the model utilized in this study exhibits sensitivity in estimating interaction strengths (IS) among species with exceedingly small body masses (approximately $1 x 10 ^{-15}$ kg), as evidenced by a broad interquartile range. Additionally, some of the observed results may stem from the fact that the applied methodology does not incorporate empirically measured biomasses for species. For example, we would have expected a relatively high impact on community stability when removing the Antarctic krill Euphausia superba, as it is one of the most important prey species in the food web based on field observations [@Atkinson2019]. Moreover, it is one of the species most affected by warming and its population has already declined as a result of productivity changes and increased access by predators [@Atkinson2004; @Kawaguchi2009]. Although our results did not show a significant impact on the stability of the Weddell Sea food web when krill went extinct, we were able to frame it as a key species in terms of its weighted and unweighted characteristics due to its high mean IS, mid trophic level and high number of interactions.

### 2023-12-18 Manuscript/Manuscript.tex
Incorporating editors suggs

The high Antarctic Weddell Sea shelf is situated between 74 and 78ºS,
stretching approximately 450 km from East to West (Figure 1). Water
depth varies between 200 and 500 meters, and shallower areas are covered
by continental ice, which forms the coastline along the eastern and
southern parts of the Weddell Sea. The shelf area contains a complex
three-dimensional benthic habitat with large benthic biomasses,
intermediate to high diversity in comparison to benthic boreal
communities and a spatially patchy distribution of organisms
\citep{Dayton1990, Teixido2002}.

@@

The Weddell Sea food web was retrieved from the GlobAL daTabasE of
traits and food Web Architecture (GATEWAy, version 1.0) of the German
Centre for Integrative Biodiversity Research (iDiv) Halle-Jena-Leipzig
\citep{Brose2018}. In addition to predator-prey interactions, the
database contains information on other biological data such as mean body
mass and movement type for each species in the food web. Furthermore, it
incorporates information about the interaction itself, such as the
dimension of the predator search space (2 or 3 dimensions). In its
current form, the Weddell Sea food web comprises 490 species and 16041
predator-prey interactions and thus constitutes one of the most resolved
food webs constructed to date \citep{Jacob2011}.

Complex empirical food webs are compilations of species and their
potential feeding interactions within a given area and time period. The
species occurrence data to construct the Weddell Sea food web
(Antarctica) was sampled between 1983 and 2005, and hence the Weddell
Sea food web is representative for this time period. The diet
composition for each species was observed between 2001 and 2004 through
a combination of field observations and stomach content analyses
(further details can be found in \citet{Jacob2005}). Some species, such
as benthic grazers and suspension feeders had poor taxonomic resolution
of their prey, therefore laboratory information about their size,
behavior, and stable isotope signatures was used to infer their feeding
habits \citep{Jacob2011}. As the data used to construct the Weddell Sea
food web is sampled during the summer season due to the inaccessibility
of ecological field sampling in Antarctica during the winter, the
Weddell Sea food web is representative of the summer season. Moreover,
the species in the food web are representative of adult specimens, and
were not categorized into ontogenetic life stages (i.e., larvae,
juveniles, and adults). As a result, ontogenetic diet shifts cannot be
addressed with this food web. The average body masses of species were
either directly measured \citep{Jacob2005} or taken from published
accounts for marine mammals and seabirds \citep{Brose2006a}. Due to the
static nature of the Weddell Sea food web and its low temporal
resolution, it is not possible to assess phenological mismatches nor
species temporal turnover. Because of these methodological limitations,
the nature of the Weddell Sea food web data may likely lead to
overestimations of some trophic interactions and network properties
(connectance), compared to realised networks at any given point in time
and space. The spatial extent of the food web is described in the Study
area section (Figure 1).

@@

Here it is important to note that we observed an important variability
when estimating each IS. The interactions that showed the greatest
variability, represented by a wide interquartile range, were those where
consumer and resource body masses were very small (approximately 1e-15
kg). The frequency distribution of the interquartile range is presented
in the Supplementary Material (Figure S1).

@@

Another source of uncertainty stems from the parameters and data used to
estimate species IS. While we accounted for variability from estimated
resource density and consumer search rate \citep{Pawar2012}, we did not
consider potential variability arising from changes in species' body
mass. Indeed, one could estimate the interaction strength based on body
mass distributions rather than a single empirical value. Future research
should consider this source of variability since it could alter food web
stability outcomes
\citep{Brose2006a, Allesina2015, Gross2009, Landi2018}. Concerning body
mass, the model utilized in this study exhibits sensitivity in
estimating interaction strengths (IS) among species with exceedingly
small body masses (approximately 1e-15 kg), as evidenced by a broad
interquartile range. Additionally, some of the observed results may stem
from the fact that the applied methodology does not incorporate
empirically measured biomasses for species. For example, we would have
expected a relatively high impact on community stability when removing
the Antarctic krill Euphausia superba, as it is one of the most
important prey species in the food web based on field observations
\citep{Atkinson2019}. Moreover, it is one of the species most affected
by warming and its population has already declined as a result of
productivity changes and increased access by predators
\citep{Atkinson2004, Kawaguchi2009}. Although our results did not show a
significant impact on the stability of the Weddell Sea food web when
krill went extinct, we were able to frame it as a key species in terms
of its weighted and unweighted characteristics due to its high mean IS,
mid trophic level and high number of interactions.

>>>>

The high Antarctic Weddell Sea shelf, our study area, is situated
between 74 and 78ºS, stretching approximately 450 km from East to West
(Figure 1). Water depth varies between 200 and 500 meters, and shallower
areas are covered by continental ice, which forms the coastline along
the eastern and southern parts of the Weddell Sea. The shelf area
contains a complex three-dimensional benthic habitat with large benthic
biomasses, intermediate to high diversity in comparison to benthic
boreal communities and a spatially patchy distribution of organisms
\citep{Dayton1990, Teixido2002}.

@@

The Weddell Sea food web was retrieved from the GlobAL daTabasE of
traits and food Web Architecture (GATEWAy, version 1.0) of the German
Centre for Integrative Biodiversity Research (iDiv) Halle-Jena-Leipzig
\citep{Brose2018}. In addition to predator-prey interactions, the
database contains information on other biological data: mean body mass
and length, and movement and metabolic type for each species in the food
web. Furthermore, it incorporates information about the interaction
itself: type (predacious, herbivorous and detritivorous) and dimension
(2 or 3 dimensions according to the predator search space). In its
current form, the Weddell Sea food web comprises 490 species and 16041
predator-prey interactions and thus constitutes one of the most resolved
food webs constructed to date \citep{Jacob2011}.

Complex empirical food webs are compilations of species and their
potential feeding interactions within a given area and time period. The
species occurrence data to construct the Weddell Sea food web was
sampled between 1983 and 2005, and hence the Weddell Sea food web is
representative for this time period. The diet composition for each
species was observed between 2001 and 2004 through a combination of
field observations and stomach content analyses (further details can be
found in \citet{Jacob2005}). Some species, such as benthic grazers and
suspension feeders had poor taxonomic resolution of their prey,
therefore laboratory information about their size, behavior, and stable
isotope signatures was used to infer their feeding habits
\citep{Jacob2011}. As the data used to construct the Weddell Sea food
web is sampled during the summer season due to the inaccessibility of
ecological field sampling in Antarctica during the winter, the Weddell
Sea food web is representative of the summer season. Moreover, the
species in the food web are representative of adult specimens, and were
not categorized into ontogenetic life stages (i.e., larvae, juveniles,
and adults). As a result, ontogenetic diet shifts cannot be addressed
with this food web. The average body masses of species were either
directly measured \citep{Jacob2005} or taken from published accounts for
marine mammals and seabirds \citep{Brose2006a}. Due to the static nature
of the Weddell Sea food web and its low temporal resolution, it is not
possible to assess phenological mismatches nor species temporal
turnover. Because of these methodological limitations, the nature of the
Weddell Sea food web data may likely lead to overestimations of some
trophic interactions and network properties (connectance), compared to
realised networks at any given point in time and space. The spatial
extent of the food web is described in the Study area section (Figure
1).

@@

Here it is important to note that we observed an important variability
when estimating each IS. The interactions that showed the greatest
variability, represented by a wide interquartile range, were those where
consumer and resource body masses were very small (approximately
\(1 x 10 ^{-15}\) kg). The frequency distribution of the interquartile
range is presented in the Supplementary Material (Figure S1).

@@

Another source of uncertainty stems from the parameters and data used to
estimate species IS. While we accounted for variability from estimated
resource density and consumer search rate \citep{Pawar2012}, we did not
consider potential variability arising from changes in species' body
mass. Indeed, one could estimate the interaction strength based on body
mass distributions rather than a single empirical value. Future research
should consider this source of variability since it could alter food web
stability outcomes
\citep{Brose2006a, Allesina2015, Gross2009, Landi2018}. Concerning body
mass, the model utilized in this study exhibits sensitivity in
estimating interaction strengths (IS) among species with exceedingly
small body masses (approximately \(1 x 10 ^{-15}\) kg), as evidenced by
a broad interquartile range. Additionally, some of the observed results
may stem from the fact that the applied methodology does not incorporate
empirically measured biomasses for species. For example, we would have
expected a relatively high impact on community stability when removing
the Antarctic krill Euphausia superba, as it is one of the most
important prey species in the food web based on field observations
\citep{Atkinson2019}. Moreover, it is one of the species most affected
by warming and its population has already declined as a result of
productivity changes and increased access by predators
\citep{Atkinson2004, Kawaguchi2009}. Although our results did not show a
significant impact on the stability of the Weddell Sea food web when
krill went extinct, we were able to frame it as a key species in terms
of its weighted and unweighted characteristics due to its high mean IS,
mid trophic level and high number of interactions.
